---
title: "Bayesian Structural Time Series"
author: "Isaac Vernon"
date: "6/15/2021"
output: html_document
---


Example from StitchFix, using new data

THE DATA IS SETUP TO FORECAST 2017, 2018, 2019

```{r}
library(bsts)
library(dplyr)
library(Metrics)

timeseries = read.csv("/Users/Isaac/Documents/GitHub/W210-Capstone/models/Time Series/QuarterlyFredWithItems.csv")

```

### GDPC1

```{r}
# 259 - 2012
# 263 - 2013
# 267 - 2014
# 271 - 2015
# 275 - 2016
# 279 - 2017
# 283 - 2018
# 287 - 2019
# 291 - 2020
# 295 - 2021
y = timeseries$GDPC1_PCH[1:255]
actuals = timeseries$GDPC1_PCH[256:259]#291]
ss <- AddLocalLinearTrend(list(), y)
ss <- AddSeasonal(ss, y, nseasons = 4)
bsts.model <- bsts(y, state.specification = ss, niter = 500, ping=0, seed=2016)

```

```{r}
burn <- SuggestBurn(0.1, bsts.model)
p <- predict.bsts(bsts.model, horizon = 4, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
#aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
aggpred = sum(preds)
rmse(actuals, preds)
#aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
aggactuals = sum(actuals)
rmse(aggactuals, aggpred)
aggpred
```

```{r}
### Actual versus predicted
d2 <- data.frame(
    # fitted values and predictions
    c(as.numeric(-colMeans(bsts.model$one.step.prediction.errors[-(1:burn),])+y),  
    as.numeric(p$mean)),
    # actual data and dates 
    as.numeric(timeseries$GDPC1_PCH[1:291]))
names(d2) <- c("Fitted", "Actual")

### MAPE (mean absolute percentage error)
MAPE <- d2 %>% summarise(MAPE=mean(abs(Actual-Fitted)/Actual))


### Plot actual versus predicted with credible intervals for the holdout period
plot(d2$Fitted, type="l")
plot(d2$Actual, type="l")
```


Try the Variable Selection Time Series (Shifts each GDP value a quarter forward -- to remove present time dependencies)
```{r}
timeseries = read.csv("/Users/Isaac/Documents/GitHub/W210-Capstone/models/Time Series/QuarterlyFredWithItems.csv")
# 235 - 2005
# 239 - 2006
# 243 - 2007
# 247 - 2008
# 251 - 2009
# 255 - 2010

# 259 - 2011
# 263 - 2012
# 267 - 2013
# 271 - 2014
# 275 - 2015
# 279 - 2016
# 283 - 2017
# 287 - 2018
# 291 - 2019
# 295 - 2020

# 299 - 2021
timeseries = transform(timeseries, GDPC1_PCH=lead(GDPC1_PCH))
timeseries2 = timeseries[180:291, 2:18]
timeseries2$pred_average = rep(datas$pred_average[1:28]/4, each=4)
timeseries2$conf = rep(datas$FORECASTER_CONF[1:28]/4, each=4)

newdata = timeseries[292:296, 3:18]
newdata$pred_average = c(rep(datas$pred_average[29]/4, each = 4), datas$pred_average[30]/4)
newdata$conf = c(rep(datas$FORECASTER_CONF[29]/4, each = 4), datas$FORECASTER_CONF[30]/4)
actuals = timeseries[292:296, 2]

ss <- AddLocalLinearTrend(list(), timeseries2$GDPC1_PCH)
ss <- AddSeasonal(ss, timeseries2$GDPC1_PCH, nseasons = 4)
bsts.reg <- bsts(GDPC1_PCH ~ ., state.specification = ss, data =
                timeseries2, niter = 500, ping=0, seed=2016)

p <- predict.bsts(bsts.reg, newdata=newdata, horizon = 5, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
#aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
aggpred = sum(preds)
#rmse(actuals, preds)
#aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
#rmse(aggactuals, aggpred)
#h = append(h, aggpred)
#l = array(aggpred)
#l = append(l, aggpred)
```


```{r}
plot(preds, type="l")
plot(timeseries$GDPC1_PCH[280:291], type="l")
```


```{r}
library(reshape2)
library(ggplot2)

burn <- SuggestBurn(0.1, bsts.reg)
### Helper function to get the positive mean of a vector
PositiveMean <- function(b) {
  b <- b[abs(b) > 0]
  if (length(b) > 0) 
    return(mean(b))
  return(0)
}

### Get the average coefficients when variables were selected (non-zero slopes)
coeff <- data.frame(melt(apply(bsts.reg$coefficients[-(1:burn),], 2, PositiveMean)))
coeff$Variable <- as.character(row.names(coeff))
ggplot(data=coeff, aes(x=Variable, y=value)) + 
  geom_bar(stat="identity", position="identity") + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  xlab("") + ylab("") + ggtitle("Average coefficients")
```

```{r}
components.withreg <- cbind.data.frame(
  colMeans(bsts.reg$state.contributions[-(1:burn),"trend",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"seasonal.4.1",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"regression",]),
  as.Date(time(timeseries$DATE[149:279])))  
names(components.withreg) <- c("Trend", "Seasonality", "Regression", "Date")
components.withreg <- melt(components.withreg, id.vars="Date")
names(components.withreg) <- c("Date", "Component", "Value")

ggplot(data=components.withreg, aes(x=Date, y=Value)) + geom_line() + 
  theme_bw() + theme(legend.title = element_blank()) + ylab("") + xlab("") + 
  facet_grid(Component ~ ., scales="free") + guides(colour=FALSE) + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0))
```


### Unemployment Rate


```{r}
timeseries = read.csv("/Users/Isaac/Documents/GitHub/W210-Capstone/data/CleanData/QuarterlyFredWithItems.csv")


y = timeseries$UNRATE[1:279]
actuals = timeseries$UNRATE[280:291]
ss <- AddLocalLinearTrend(list(), y)
ss <- AddSeasonal(ss, y, nseasons = 4)
bsts.model <- bsts(y, state.specification = ss, niter = 500, ping=0, seed=2016)

```

```{r}
burn <- SuggestBurn(0.1, bsts.model)
p <- predict.bsts(bsts.model, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```

```{r}
### Actual versus predicted
d2 <- data.frame(
    # fitted values and predictions
    c(as.numeric(-colMeans(bsts.model$one.step.prediction.errors[-(1:burn),])+y),  
    as.numeric(p$mean)),
    # actual data and dates 
    as.numeric(timeseries$UNRATE[1:291]))
names(d2) <- c("Fitted", "Actual")

### MAPE (mean absolute percentage error)
MAPE <- d2 %>% summarise(MAPE=mean(abs(Actual-Fitted)/Actual))


### Plot actual versus predicted with credible intervals for the holdout period
plot(d2$Fitted, type="l")
plot(d2$Actual, type="l")
```


Try the Variable Selection Time Series (Shifts each GDP value a quarter forward -- to remove present time dependencies)
```{r}
timeseries = transform(timeseries, UNRATE=lead(UNRATE))
timeseries2 = timeseries[149:279, 2:18]

newdata = timeseries[280:291, c(2, 3, 4, 5, 6, 7, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18)]
actuals = timeseries[280:291, 8]


ss <- AddLocalLinearTrend(list(), timeseries2$UNRATE)
ss <- AddSeasonal(ss, timeseries2$UNRATE, nseasons = 4)
bsts.reg <- bsts(UNRATE ~ ., state.specification = ss, data =
                timeseries2, niter = 500, ping=0, seed=2016)

p <- predict.bsts(bsts.reg, newdata=newdata, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```


```{r}
plot(preds, type="l")
plot(timeseries$UNRATE[280:291], type="l")
```


```{r}
library(reshape2)
library(ggplot2)

burn <- SuggestBurn(0.1, bsts.reg)
### Helper function to get the positive mean of a vector
PositiveMean <- function(b) {
  b <- b[abs(b) > 0]
  if (length(b) > 0) 
    return(mean(b))
  return(0)
}

### Get the average coefficients when variables were selected (non-zero slopes)
coeff <- data.frame(melt(apply(bsts.reg$coefficients[-(1:burn),], 2, PositiveMean)))
coeff$Variable <- as.character(row.names(coeff))
ggplot(data=coeff, aes(x=Variable, y=value)) + 
  geom_bar(stat="identity", position="identity") + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  xlab("") + ylab("") + ggtitle("Average coefficients")
```

```{r}
components.withreg <- cbind.data.frame(
  colMeans(bsts.reg$state.contributions[-(1:burn),"trend",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"seasonal.4.1",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"regression",]),
  as.Date(time(timeseries$DATE[149:279])))  
names(components.withreg) <- c("Trend", "Seasonality", "Regression", "Date")
components.withreg <- melt(components.withreg, id.vars="Date")
names(components.withreg) <- c("Date", "Component", "Value")

ggplot(data=components.withreg, aes(x=Date, y=Value)) + geom_line() + 
  theme_bw() + theme(legend.title = element_blank()) + ylab("") + xlab("") + 
  facet_grid(Component ~ ., scales="free") + guides(colour=FALSE) + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0))
```


### PCEPILFE


```{r}
timeseries = read.csv("/Users/Isaac/Documents/GitHub/W210-Capstone/data/CleanData/QuarterlyFredWithItems.csv")

y = timeseries$PCEPILFE[1:279]
actuals = timeseries$PCEPILFE[280:291]
ss <- AddLocalLinearTrend(list(), y)
ss <- AddSeasonal(ss, y, nseasons = 4)
bsts.model <- bsts(y, state.specification = ss, niter = 500, ping=0, seed=2016)

```

```{r}
burn <- SuggestBurn(0.1, bsts.model)
p <- predict.bsts(bsts.model, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```

```{r}
### Actual versus predicted
d2 <- data.frame(
    # fitted values and predictions
    c(as.numeric(-colMeans(bsts.model$one.step.prediction.errors[-(1:burn),])+y),  
    as.numeric(p$mean)),
    # actual data and dates 
    as.numeric(timeseries$PCEPILFE[1:291]))
names(d2) <- c("Fitted", "Actual")

### MAPE (mean absolute percentage error)
MAPE <- d2 %>% summarise(MAPE=mean(abs(Actual-Fitted)/Actual))


### Plot actual versus predicted with credible intervals for the holdout period
plot(d2$Fitted, type="l")
plot(d2$Actual, type="l")
```


Try the Variable Selection Time Series (Shifts each GDP value a quarter forward -- to remove present time dependencies)
```{r}
timeseries = transform(timeseries, PCEPILFE=lead(PCEPILFE))
timeseries2 = timeseries[149:279, 2:18]

newdata = timeseries[280:291, c(2, 3, 4, 5, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18)]
actuals = timeseries[280:291, 6]


ss <- AddLocalLinearTrend(list(), timeseries2$PCEPILFE)
ss <- AddSeasonal(ss, timeseries2$PCEPILFE, nseasons = 4)
bsts.reg <- bsts(PCEPILFE ~ ., state.specification = ss, data =
                timeseries2, niter = 500, ping=0, seed=2016)

p <- predict.bsts(bsts.reg, newdata=newdata, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```


```{r}
plot(preds, type="l")
plot(timeseries$PCEPILFE[280:291], type="l")
```


```{r}
library(reshape2)
library(ggplot2)

burn <- SuggestBurn(0.1, bsts.reg)
### Helper function to get the positive mean of a vector
PositiveMean <- function(b) {
  b <- b[abs(b) > 0]
  if (length(b) > 0) 
    return(mean(b))
  return(0)
}

### Get the average coefficients when variables were selected (non-zero slopes)
coeff <- data.frame(melt(apply(bsts.reg$coefficients[-(1:burn),], 2, PositiveMean)))
coeff$Variable <- as.character(row.names(coeff))
ggplot(data=coeff, aes(x=Variable, y=value)) + 
  geom_bar(stat="identity", position="identity") + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  xlab("") + ylab("") + ggtitle("Average coefficients")
```

```{r}
components.withreg <- cbind.data.frame(
  colMeans(bsts.reg$state.contributions[-(1:burn),"trend",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"seasonal.4.1",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"regression",]),
  as.Date(time(timeseries$DATE[149:279])))  
names(components.withreg) <- c("Trend", "Seasonality", "Regression", "Date")
components.withreg <- melt(components.withreg, id.vars="Date")
names(components.withreg) <- c("Date", "Component", "Value")

ggplot(data=components.withreg, aes(x=Date, y=Value)) + geom_line() + 
  theme_bw() + theme(legend.title = element_blank()) + ylab("") + xlab("") + 
  facet_grid(Component ~ ., scales="free") + guides(colour=FALSE) + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0))
```


### CPIFESL


```{r}
timeseries = read.csv("/Users/Isaac/Documents/GitHub/W210-Capstone/data/CleanData/QuarterlyFredWithItems.csv")


y = timeseries$CPILFESL[1:279]
actuals = timeseries$CPILFESL[280:291]
ss <- AddLocalLinearTrend(list(), y)
ss <- AddSeasonal(ss, y, nseasons = 4)
bsts.model <- bsts(y, state.specification = ss, niter = 500, ping=0, seed=2016)

```

```{r}
burn <- SuggestBurn(0.1, bsts.model)
p <- predict.bsts(bsts.model, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```

```{r}
### Actual versus predicted
d2 <- data.frame(
    # fitted values and predictions
    c(as.numeric(-colMeans(bsts.model$one.step.prediction.errors[-(1:burn),])+y),  
    as.numeric(p$mean)),
    # actual data and dates 
    as.numeric(timeseries$CPILFESL[1:291]))
names(d2) <- c("Fitted", "Actual")

### MAPE (mean absolute percentage error)
MAPE <- d2 %>% summarise(MAPE=mean(abs(Actual-Fitted)/Actual))


### Plot actual versus predicted with credible intervals for the holdout period
plot(d2$Fitted, type="l")
plot(d2$Actual, type="l")
```


Try the Variable Selection Time Series (Shifts each GDP value a quarter forward -- to remove present time dependencies)
```{r}
timeseries = transform(timeseries, CPILFESL=lead(CPILFESL))
timeseries2 = timeseries[149:279, 2:18]

newdata = timeseries[280:291, c(2, 3, 4, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18)]
actuals = timeseries[280:291, 5]


ss <- AddLocalLinearTrend(list(), timeseries2$CPILFESL)
ss <- AddSeasonal(ss, timeseries2$CPILFESL, nseasons = 4)
bsts.reg <- bsts(CPILFESL ~ ., state.specification = ss, data =
                timeseries2, niter = 500, ping=0, seed=2016)

p <- predict.bsts(bsts.reg, newdata=newdata, horizon = 12, burn = burn, quantiles = c(.025, .975))

preds = as.numeric(p$mean)
aggpred = c(mean(preds[1:4]), mean(preds[5:8]), mean(preds[9:12]))
rmse(actuals, preds)
aggactuals = c(mean(actuals[1:4]), mean(actuals[5:8]), mean(actuals[9:12]))
rmse(aggactuals, aggpred)
```


```{r}
plot(preds, type="l")
plot(timeseries$CPILFESL[280:291], type="l")
```


```{r}
library(reshape2)
library(ggplot2)

burn <- SuggestBurn(0.1, bsts.reg)
### Helper function to get the positive mean of a vector
PositiveMean <- function(b) {
  b <- b[abs(b) > 0]
  if (length(b) > 0) 
    return(mean(b))
  return(0)
}

### Get the average coefficients when variables were selected (non-zero slopes)
coeff <- data.frame(melt(apply(bsts.reg$coefficients[-(1:burn),], 2, PositiveMean)))
coeff$Variable <- as.character(row.names(coeff))
ggplot(data=coeff, aes(x=Variable, y=value)) + 
  geom_bar(stat="identity", position="identity") + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  xlab("") + ylab("") + ggtitle("Average coefficients")
```

```{r}
components.withreg <- cbind.data.frame(
  colMeans(bsts.reg$state.contributions[-(1:burn),"trend",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"seasonal.4.1",]),
  colMeans(bsts.reg$state.contributions[-(1:burn),"regression",]),
  as.Date(time(timeseries$DATE[149:279])))  
names(components.withreg) <- c("Trend", "Seasonality", "Regression", "Date")
components.withreg <- melt(components.withreg, id.vars="Date")
names(components.withreg) <- c("Date", "Component", "Value")

ggplot(data=components.withreg, aes(x=Date, y=Value)) + geom_line() + 
  theme_bw() + theme(legend.title = element_blank()) + ylab("") + xlab("") + 
  facet_grid(Component ~ ., scales="free") + guides(colour=FALSE) + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0))
```